---
title: "Web scraping and basic data analysis"
author: "Greinald Pappa (12046752), Arjol Pançi(12347494)"
date: "Apr. 13, 2025 (Assignment_1.Rmd)"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: true
    citation_package: natbib
    latex_engine: xelatex
bibliography: literature.bib
biblio-style: humannat
include-before: \pagebreak
editor_options:
  markdown:
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
# Configure knitr options and suppress warnings/messages
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required libraries
library(tidyverse)      # For data manipulation and visualization
library(tidytext)       # For handling textual data
library(wordcloud)      # For generating word clouds
library(tm)             # For text mining preprocessing
library(ggplot2)        # For customized plots
library(RColorBrewer)   # For aesthetic color palettes in word clouds

# Set system locale to German for correct date parsing (if needed)
Sys.setlocale("LC_TIME", "de_DE.UTF-8")

```


## Introduction

This report investigates how technology impacts student life by analyzing survey data. We focus on demographic patterns, digital usage, text responses, and statistical associations. The goal is to draw both descriptive and inferential conclusions on how students interact with technology in their daily academic routines.

## Data Preparation

```{r}
# Load survey data from CSV file
survey_data <- read.csv('survey_data.csv', stringsAsFactors = FALSE)

# Convert key categorical variables into factor types for analysis
survey_data$Gender <- factor(survey_data$Gender,
                             levels = c("Woman", "Man", "Non-binary", "Prefer not to say"))
survey_data$Age <- factor(survey_data$Age,
                          levels = c("18-21", "22-25", "26-30", "30+"))

# Convert digital usage column from character to numeric
survey_data$Hours_Digital_Use <- as.numeric(survey_data$Hours_Digital_Use)
```

We began by loading the data and converting relevant columns to the correct types. Factor levels were manually specified to ensure consistent ordering in plots and analyses.

## Exploratory Data Analysis

We begin by examining basic demographics and behavioral patterns among students.

```{r}
# Plot gender distribution
ggplot(survey_data, aes(x = Gender)) +
  geom_bar(fill = '#4682B4', color = 'black') +
  theme_minimal(base_size = 14) +
  labs(title = "Gender Distribution of Respondents", x = "Gender", y = "Count")
```

```{r}
# Plot age group distribution
ggplot(survey_data, aes(x = Age)) +
  geom_bar(fill = '#2E8B57', color = 'black') +
  theme_minimal(base_size = 14) +
  labs(title = "Age Distribution of Participants", x = "Age Group", y = "Count")
```

```{r}
# Count and display academic program frequencies
program_counts <- survey_data %>%
  count(Academic_Program, sort = TRUE)

# Display as a table
knitr::kable(program_counts, caption = "Frequency of Academic Programs")
```

```{r}
# Histogram of daily device usage in hours
ggplot(survey_data, aes(x = Hours_Digital_Use)) +
  geom_histogram(binwidth = 1, fill = '#FF7F50', color = 'black') +
  theme_minimal(base_size = 14) +
  labs(title = "Daily Use of Digital Devices", x = "Hours per Day", y = "Number of Students")
```

These plots provide insight into the student population, highlighting usage patterns and academic program representation.

## Textual Analysis

We now analyze open-ended responses to questions about the academic benefits of technology and the effects of social media.

```{r}
# Preprocess and clean text related to academic improvement
academics_corpus <- Corpus(VectorSource(survey_data$Tech_Improved_Academics)) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeWords, stopwords("en"))

# Generate a word cloud for academic improvements
wordcloud(academics_corpus, max.words = 100, random.order = FALSE,
          colors = brewer.pal(8, "Dark2"))
```

```{r}
# Preprocess and clean text related to social media impact
socialmedia_corpus <- Corpus(VectorSource(survey_data$Social_Media_Impact)) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeWords, stopwords("en"))

# Generate a word cloud for social media impact
wordcloud(socialmedia_corpus, max.words = 100, random.order = FALSE,
          colors = brewer.pal(8, "Set3"))
```

The word clouds provide a visual summary of key terms in student feedback. Words like "helped", "easier", and "distraction" give insight into student perspectives.

## Descriptive Inference

We compute key summary statistics to better understand the range and average of digital usage.

```{r}
# Calculate mean, median, SD, min, and max for daily usage
summary_hours <- survey_data %>% summarise(
  Mean = mean(Hours_Digital_Use, na.rm=TRUE),
  Median = median(Hours_Digital_Use, na.rm=TRUE),
  SD = sd(Hours_Digital_Use, na.rm=TRUE),
  Min = min(Hours_Digital_Use, na.rm=TRUE),
  Max = max(Hours_Digital_Use, na.rm=TRUE)
)

# Display the summary table
knitr::kable(summary_hours, caption = "Summary of Hours Spent on Digital Devices")
```

This table confirms that while most students use digital devices extensively, there’s considerable variation between individuals.

## Analytic Inference

We now test for significant differences and associations using statistical methods.

```{r}
# Filter for binary gender responses and remove NA values
usages_gender <- survey_data %>%
  filter(Gender %in% c("Man", "Woman")) %>%
  drop_na(Hours_Digital_Use)

# Perform a t-test to compare usage between men and women
usage_test <- t.test(Hours_Digital_Use ~ Gender, data = usages_gender)

# Show t-test results
usage_test
```

The t-test checks whether gender is a significant factor in digital usage. A non-significant p-value would suggest no strong difference.

```{r}
# Create contingency table between age and academic program
age_program_table <- table(survey_data$Age, survey_data$Academic_Program)

# Perform chi-square test of independence
age_program_chisq <- chisq.test(age_program_table)

# Show chi-square test results
age_program_chisq
```

The chi-square test evaluates whether age group and academic program are independent. A significant result implies a relationship between them.

## Conclusion

This report shows that digital technology is central to student life, with high daily usage and mixed impacts on academic performance. Gender does not appear to significantly influence device use, but academic program and age group are associated. Open-ended responses show that students acknowledge both the benefits and drawbacks of technology in their academic experience. These insights can inform future educational strategies that consider how students use and perceive technology in learning environments.
```

